
\documentclass[11pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[abbr]{harvard}
\usepackage{amssymb}
\usepackage{setspace,graphics,epsfig,amsmath,rotating,amsfonts,mathpazo}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.50.0.2960}
%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
%TCIDATA{BibliographyScheme=BibTeX}
%TCIDATA{Created=Thursday, September 11, 2008 15:11:56}
%TCIDATA{LastRevised=Wednesday, November 20, 2013 16:01:51}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{<META NAME="DocumentShell" CONTENT="Articles\SW\article.egan">}

\topmargin=0 in \headheight=0in \headsep=0in \topskip=0in \textheight=9in \oddsidemargin=0in \evensidemargin=0in \textwidth=6.5in
\input{tcilatex}
\begin{document}


New York University

Wilf Family Department of Politics

Fall 2013

\begin{center}
{\large \textbf{Quantitative Research in Political\ Science I}}

Professor Patrick Egan

\bigskip

\textbf{PROBLEM\ SET 8: Due Monday, November 25 at beginning of class.}
\end{center}

\textit{A reminder: you may work with others in the class on this problem
set, and you are in fact encouraged to do so. \ However, the work you hand
in must be your own. \ Your work must be word-processed in order for you to
receive credit for the assignment.}

\begin{enumerate}
\item Show (using matrix notation) that by construction in OLS 
\begin{equation*}
\mathbf{i}^{\prime }\widehat{\mathbf{u}}=0.
\end{equation*}

\item Show (using matrix notation) that by construction in OLS for any $N$ x
1 vector of observations of the $k$'th regressor $\mathbf{x}_{k}$, it is the
case that%
\begin{equation*}
cov\left( \mathbf{x}_{k},\widehat{\mathbf{u}}\right) =0.
\end{equation*}%
\bigskip

For the rest of this assignment, use the \textit{\ counties.dta} dataset
found on our class Blackboard site. It comes from the \textit{City and
County Data Book }produced by the U.S.\ Census Bureau. \ Provide appropriate
Stata output as an attachment to your assignment.

\item Analyze the following questions using Stata, but answer them with a
few sentences in plain English. \ 

\begin{enumerate}
\item Controlling for household income, is the percentage of a county's
residents who have college degrees associated with crime rates? \ How so?

\item Construct a plot displaying the relationship between $\widehat{\mathit{%
crimerate}}$ and \textit{college}. \ To do this, write a loop that generates
the value of $\widehat{\mathit{crimerate}}$ \ when \textit{college }is equal
to 5, 7,...25. \ (In doing so, you may find it helpful to use the
coefficients saved by the \texttt{regress} command in the vector \texttt{e(b)%
}.)\ Then plot these connected points.

\item You are interested in how well this model applies to counties in New
York State. \ Construct a plot showing at a glance that Tompkins County, has
a crime rate much lower than predicted by a regression of \textit{crimerate}
on \textit{college }among counties in New York State.

\item Look up Tompkins County on the Internet. \ Why might it be an outlier?
\ Does this suggest an additional control variable?\bigskip
\end{enumerate}

The following two questions require that you use two techniques not yet
covered in class: OLS with indicator variables and with interaction terms. \
Many of you are already familiar with these techniques, and Andrew will be
covering them in lab on this Friday. \ I encourage you to consult the cited
sections of Wooldridge if needed. \ We'll cover them in detail next week in
class.\bigskip

\item Is unemployment higher in counties located in the South than in those
located outside the South?

\begin{enumerate}
\item Answer this question first with a $t$-test to compare group means.

\item Now answer this question by running a bivariate specification
regressing \textit{unemprate} on the indicator variable \textit{south. }If
you are unsure about how to interpret the coefficient on \textit{south }in
this context, consult Wooldridge pp. 225-231.

\item Identify two important similarities found between the two analyses. \ 

\item We required a lot more assumptions to perform the second analysis than
the first. \ In a few sentences, reflect upon why they are unnecessary to
answer the question posed at (4) above.\bigskip
\end{enumerate}

\item Is the association between poverty and crime by county stronger in the
South or outside the South?

\begin{enumerate}
\item Answer this question first by running two separate regressions.

\item Now create a term \textit{povxsouth} that incorporates the interaction
between \textit{south} and \textit{povrate}. \ Run the proper analysis that
includes the interaction term. \ If needed, consult Wooldridge pp. 238-243.

\item Identify two important similarities found between the estimates
generated by the two analyses.

\item Construct a plot with two lines: one showing $\widehat{\mathit{%
crimerate}}$\textit{\ }by \textit{povrate }among counties in the South, and
the other showing $\widehat{\mathit{crimerate}}$\textit{\ \ }by\textit{\
povrate }among non-Southern counties.\bigskip
\end{enumerate}

\item Perform the necessary analyses to assess the accuracy of the following
statements. \ In a few sentences, discuss each of your findings.

\begin{enumerate}
\item The bivariate relationship between poverty and crime is weaker where
local government spending is higher.

\item The bivariate relationship between unemployment and crime is stronger
in densely populated counties.

\item Federal spending per capita plays a greater role in reducing poverty
than local spending per capita, controlling for variables that may confound
these relationships.\bigskip
\end{enumerate}

\item Pretend that you are a Marxist-Socialist who wants to show that crime
is caused by poverty. \ To do so, you are hell-bent on finding the
combination of regressors (including \textit{povrate}) that results in a
very large coefficient on \textit{povrate} when \textit{crimerate }is
regressed upon these regressors\textit{.}

\begin{enumerate}
\item Using the proper Stata commands, choose 90 percent of your
observations at random and set them aside. \ Do not include these
observations in the analysis.

\item Find the combination of variables that yields the highest possible
coefficient on \textit{povrate }as a predictor of \textit{crimerate }that
you can obtain. \ If you're feeling creative, you should be able to write a
Stata loop that accomplishes this, but that is not necessary. \ Just have
fun with it. \ [ONE MORE QUESTION ON NEXT PAGE]

\item Once you've arrived at a specification, note the results. \ Now, take
the rest of your dataset out of \textquotedblleft cold
storage.\textquotedblright\ \ How well does your model do with these other
data? \ What is the lesson learned? \ 
\end{enumerate}
\end{enumerate}

\end{document}
