---
title: "Lecture 16"
subtitle: "Quantitative Political Science"
author: "Prof. Bisbee"
institute: "Vanderbilt University"
date: "Lecture Date: 2023/11/02\n Slides Updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    # self_contained: true
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css:
      - default
      - css/lexis.css
      - css/lexis-fonts.css
    #seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "17:9"

---

```{css,echo = F}
.small .remark-code { /*Change made here*/
  font-size: 85% !important;
}
.tiny .remark-code { /*Change made here*/
  font-size: 50% !important;
}
```

```{r,include=F}
set.seed(123)
options(width=60)
knitr::opts_chunk$set(fig.align='center',fig.width=9,fig.height=5)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
require(tidyverse)
require(ggpubr)
```

# Agenda

1. Variance of OLS estimators

2. 

---

# Recap

- We went over 4 assumptions to characterize the bias of our OLS estimators

1. Relationship between $x$ and $y$ is **linear in its parameters**

2. $x$ and $y$ are drawn from a random sample, making them **i.i.d.**

3. $VAR(X) \neq 0$

4. $E(u|x) = 0$

- With these, we demonstrated that $\hat{\beta}_0$ and $\hat{\beta}_1$ are **unbiased** for $\beta_0$ and $\beta_1$

---

# Sampling Distributions

- $\hat{\beta}_0$ and $\hat{\beta}_1$ are statistics, just like $\bar{Y}$

- When we evaluate statistics, we care about both their **bias** (last class) and their **variance**

  - How far can we expect them to be from their true value (i.e., the population parameter) on average?
  
- In the univariate case, we were interested in the **sampling distribution** of $\bar{Y}$

- Here, we are also interested in the sampling distributions of $\hat{\beta}_0$ and $\hat{\beta}_1$

---

# Variance

- We already know the means of $\hat{\beta}_0$ and $\hat{\beta}_1$: they are $\beta_0$ and $\beta_1$ (from last class)

- To compute the **variances** of $\hat{\beta}_0$ and $\hat{\beta}_1$, we need a **fifth assumption**

Assumption 5: $VAR(u|x) = \sigma^2$

  - The error has the **same variance** regardless of the value of $x$
  
- This is known as **homoskedasticity**

- If this fails, we have **heteroskedasticity**

---

# Variance

```{r,message=F,warning =F}
require(tidyverse)
set.seed(123)
X <- rnorm(500,mean = 12,sd = 4)
Y <- rnorm(500,mean = X,sd = X/3)
Y2 <- rnorm(500,mean = X,sd = 2)

p <- data.frame(X = X,Heteroskedastic = Y,Homoskedastic = Y2) %>%
  gather(outcome,value,-X) %>%
  mutate(outcome = factor(outcome,levels = c('Homoskedastic','Heteroskedastic'))) %>%
  ggplot(aes(x = X,y = value)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  facet_wrap(~outcome)
```

---

# Variance

```{r,message=F,warning =F}
p
```

---

# Variance

- Note the difference between Assumptions 4 and 5!

  - $E(u|x) = 0$
  
  - $VAR(u|x) = \sigma^2$
  
- We don't need assumption 5 for unbiasedness, but we do for variance!

- What is $VAR(y|x)$?

---

# Variance

$$
\begin{aligned}
VAR(y|x) &= VAR(\beta_0 + \beta_1 x + u | x)\\
&= VAR(\beta_0 | x) + VAR(\beta_1 x | x) + VAR(u | x)\\
&= 0 + 0 + \sigma^2\\
&= \sigma^2
\end{aligned}
$$
- What is $\sigma^2$?

--

- It is a measure of the **extent to which unexplained factors are affecting $y$**

  - These factors are not related to $x$ (from assumption 4)
  
  - These factors are constant regardless of $x$ (from assumption 5)
  
  - When $\sigma^2$ is big, it means that other factors explain a lot of variation in $y$ beyond just $x$
  
  - When $\sigma^2$ is small, it means that $x$ explains a lot of variation in $y$
  
- Note that $\sigma^2$ is a **parameter**, something that exists in the population
  
---

# Sampling Variance

- Since $\sigma^2$ is a parameter, we want to estimate it in our sample

- First, $\hat{\beta}_1$

$$
\begin{aligned}
\hat{\beta}_1 &= \beta_1 + \frac{\sum(x_i - \bar{x})u_i}{SST_x} \\
VAR(\hat{\beta}_1) &= 
\end{aligned}
$$


```{r,message=F,echo=F,warning=F,results='hide'}
dir <- getwd()
type <- 'pdf'
format <- 'landscape'
f <- 'Lecture_16_slides'

# system(paste('Rscript ../NFGH/chromeprint.R',dir,type,format,f),wait = F)
```
