
\documentclass[11pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[abbr]{harvard}
\usepackage{amssymb}
\usepackage{setspace,graphics,epsfig,amsmath,rotating,amsfonts,mathpazo}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.50.0.2953}
%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
%TCIDATA{BibliographyScheme=BibTeX}
%TCIDATA{Created=Thursday, September 11, 2008 15:11:56}
%TCIDATA{LastRevised=Thursday, December 03, 2009 15:14:03}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{<META NAME="DocumentShell" CONTENT="Articles\SW\article.egan">}

\topmargin=0 in \headheight=0in \headsep=0in \topskip=0in \textheight=9in \oddsidemargin=0in \evensidemargin=0in \textwidth=6.5in
\input{tcilatex}
\begin{document}


New York University

Wilf Family Department of Politics

Fall 2009

\begin{center}
{\large \textbf{Quantitative Research in Political\ Science I}}

(G53.1150)

Professor Patrick Egan

\bigskip

\textbf{PROBLEM\ SET 8: Due Monday, December 7 at beginning of class.}
\end{center}

\textit{A reminder: you may work with others in the class on this problem
set, and you are in fact encouraged to do so. \ However, the work you hand
in must be your own. \ Your work must be word-processed in order for you to
receive credit for the assignment.}

\begin{enumerate}
\item Show (using matrix notation) that by construction in OLS (where $%
\mathbf{i}$ is a vector of ones as in Matrix Algebra Handout II, page 4), 
\begin{equation*}
\mathbf{i}^{\prime }\widehat{\mathbf{u}}=0.
\end{equation*}

\item Show (using matrix notation) that by construction in OLS for any $N$ x
1 vector of observations of the $k$'th regressor $\mathbf{x}_{k}$, it is the
case that%
\begin{equation*}
cov\left( \mathbf{x}_{k},\widehat{\mathbf{u}}\right) =0.
\end{equation*}%
\bigskip 

For the rest of this assignment, use the \textit{\ countycrime.dta} dataset
found on our class Blackboard site. (Note this week's dataset is an expanded
version of the dataset you used for Problem Set \#7; make sure you use the
right one).

\item Analyze the following questions using Stata, but answer them with a
few sentences in plain English. \ Provide appropriate Stata output as an
attachment to your assignment.

\begin{enumerate}
\item Controlling for household income, does the percentage of a county's
residents who have college degrees have any impact on crime rates? \ What is
it?

\item Construct a plot displaying the relationship between $\widehat{\mathit{%
crimerate}}$ and \textit{college}. \ To do this, write a loop that generates
the value of $\widehat{\mathit{crimerate}}$ \ when \textit{college }is equal
to 5, 7,...25. \ (In doing so, you may find it helpful to use the
coefficients saved by the \texttt{regress} command in the vector \texttt{e(b)%
}.)\ Then plot these connected points.

\item You are interested in how well this model applies to counties in New
York State. \ Construct a plot showing at a glance that Tompkins County, has
a crime rate much lower than predicted by a regression of \textit{crimerate}
on \textit{college }among counties in New York State.

\item Look up Tompkins County on the Internet. \ Why might it be an outlier?
\ What would we like to control for to see?\bigskip
\end{enumerate}

The following two questions require that you use two techniques not yet
covered in class: OLS with indicator variables and with interaction terms. \
Many of you are already familiar with these techniques. \ I encourage you to
consult the cited sections of Wooldridge if not. \ We'll cover them in
detail next week in class.\bigskip

\item Is unemployment higher in counties located in the South than in those
located outside the South?

\begin{enumerate}
\item Answer this question first with a $t$-test to compare group means.

\item Now answer this question by running a bivariate specification
regressing \textit{unemprate} on the indicator variable \textit{south. }If
you are unsure about how to interpret the coefficient on \textit{south }in
this context, consult Wooldridge pp. 225-231.

\item Identify two important similarities found between the two analyses. \ 

\item We required a lot more assumptions to perform the second analysis than
the first. \ In a few sentences, reflect upon why they are unnecessary to
answer the question posed at (4) above.\bigskip
\end{enumerate}

\item Is the association between poverty and crime by county stronger in the
South or outside the South?

\begin{enumerate}
\item Answer this question first by running two separate regressions.

\item Now create a term \textit{povxsouth} that incorporates the interaction
between \textit{south} and \textit{povrate}. \ Run the proper analysis that
includes the interaction term. \ If needed, consult Wooldridge pp. 238-243.

\item Identify two important similarities found between the estimates
generated by the two analyses.

\item Construct a plot with two lines: one showing $\widehat{\mathit{%
crimerate}}$\textit{\ }by \textit{povrate }among counties in the South, and
the other showing $\widehat{\mathit{crimerate}}$\textit{\ \ }by\textit{\
povrate }among non-Southern counties.\bigskip
\end{enumerate}

\item Pretend that you are a Marxist-Socialist who wants to show that crime
is caused by poverty. \ To do so, you are hell-bent on finding the
combination of regressors (including \textit{povrate}) that results in a
very large coefficient on \textit{povrate} when \textit{crimerate }is
regressed upon these regressors\textit{.}

\begin{enumerate}
\item Using the proper Stata commands, choose half of your observations at
random and set them aside. \ Do not include these observations in the
analysis.

\item Find the combination of variables that yields the highest possible
coefficient on \textit{povrate }that you can obtain. \ If you're feeling
creative, you should be able to write a Stata loop that accomplishes this,
but that is not necessary. \ Just have fun with it.

\item Once you've arrived at a specification, note the results. \ Now, take
the other half of your dataset out of "cold storage." \ How well does your
model do with these other data? \ What is the lesson learned? \ 
\end{enumerate}
\end{enumerate}

\end{document}
